### 2020-01-29 paper/sex_selection_spacing-ver3.tex
Add discussion of mortality to paper

% Demography abstract
\noindent 
Strong son preference is typically associated with shorter birth spacing in the absence of 
sons, but access to sex selection has the potential to reverse this pattern because each 
abortion extends spacing by six to twelve months. 
I apply an empirical method that simultaneously accounts for how sex selection 
increases both birth spacing and the likelihood of a son. 
Using the four rounds of India's National Family and Health Surveys, I show that birth 
intervals increased substantially over the last four decades, with the most significant
increases among women most likely to use sex selection.
Specifically, well-educated women with no boys now have significantly longer birth
intervals, and more male-biased sex ratios, than similar women with boys. 
Women with no education still follow the standard pattern of short average spacing when 
they have girls and little evidence of sex selection, with medium-educated women showing 
mixed results.
Furthermore, the traditional pattern of longer average spacing with sons appears to arise 
not from fewer very short birth intervals, but rather from the upper end of the distribution.

@@

India has changed significantly over the last four decades:
the economy has grown substantially,
education levels have increased for both males and females,
and the total fertility rate has fallen to the point where, at 2.2, it is just
above replacement 
\citep{Bosworth2008,Guilmoto2013,Dharmalingam2014,
International-Institute-for-Population-Sciences-IIPS2017}.

Over the same period, India has, however, also experienced a rapidly expanding 
access to prenatal sex determination.
Combined with continued strong son preference, the result has been a dramatic 
increase in the males-to-females ratio at birth
\citep{das_gupta97,Sudha1999,Arnold2002,retherford03b,jha06,Guilmoto2009a,Guilmoto2012,
Bongaarts2013,Portner2015b,Jayachandran2017}.

What the prior research on India has not appreciated is that since each abortion 
increases the interval between births by six to twelve months, the growing use of 
sex-selective abortions has the potential to increase birth spacing substantially. %
\footnote{
The increase consists of three parts.
First, starting from the abortion, the uterus needs at 
least two menstrual cycles to recover;  otherwise, the likelihood 
of spontaneous abortion increases substantially \citep{zhou00b}.
The second part is the waiting time to conception, which is between 
one and six months \citep{Wang2003}.
Finally, sex determination tests are reliable only from three months 
of gestation onwards.
}
Add to this increase that the combination of increasing female education, 
lower fertility, higher household income, and the low and declining female 
labor force participation are also likely to impact birth spacing.
Hence, the changes in birth spacing may be much more substantial in India 
than what we have observed in other countries as they developed.

Understanding the changes in birth spacing is essential for three reasons.

First, tempo effects, arising from changes in birth spacing, bias the total 
fertility rate (TFR), in the sense that the TFR becomes a more biased measure 
of cohort fertility \citep{Bongaarts1999,Ni-Bhrolchain2011}.
Hence, fertility may be, in fact, substantially higher than what is generally 
accepted based on the TFR.

Second, longer birth spacing is associated with lower mortality and 
morbidity risk \citep{Conde-Agudelo2006,Conde-Agudelo2012,Molitoris2019}.
In India, the duration to the next birth has traditionally been shorter with fewer sons, 
which likely contributed to the higher mortality and morbidity risk for girls
\citep{Whitworth2002,Bhargava2003,Rutstein2005,Bhalotra2008,Maitra2008,Makepeace2008,
Jayachandran2011,Kumar2016,Jayachandran2017a,Ghosh2018}.
Longer spacing between births then potentially lowers mortality, although there may
be a counteracting effect of multiple abortions.

Finally, birth spacing has long served as a measure of son preference, with strong
son preference associated with shorter birth intervals, the fewer sons a family has 
\citep{ben-porath76b,Leung1988}.
Now, longer birth intervals in the absence of sons, rather than an indicator of 
lower son preference, may instead arise because couples with a strong preference 
for sons use sex selection, possibly to the point where we observe a reversal of 
the traditional spacing pattern.
Families with son preference who---for one reason or another---do not use prenatal 
sex selection likely continue to have shorter birth spacing.

Here, I examine how birth spacing in India has changed over time and across 
groups with the introduction of sex selection.
I apply a competing risk hazard model that directly incorporates the effects of 
sex-selective abortions on birth intervals
\emph{and} 
the likelihood of a son, 
but which also works in cases without prenatal sex selection.
I apply the method to birth histories of Hindu women, using data from the four
India's National Family and Health Surveys (NFHS), covering the period 
1972 to 2016.

% [Hypotheses / questions]
I address four main questions.
[TK combine questions and highlight of answers here?]

First, how did average birth intervals change over time for different groups, and how do 
these changes relate to the use of sex selection?
To show how sex selection has affected birth spacing, I focus on how birth intervals have
changed for different education levels, urban or rural, and the sex composition of 
prior children.

Birth intervals have increased for all education groups over the four decades, 
with larger increases, the higher the parity and the higher the education level.
Women who are most likely to use sex selection---well-educated women with no
sons---have seen the most substantial increases in birth intervals.
[TK check this!] As a result, well-educated women now show a reversal of the traditional spacing 
pattern, with the longest intervals for women with no sons.
Those least likely to use sex selection, women with no education in rural areas, 
still follow the traditional pattern of short spacing when they have girls and 
little evidence of the use of sex selection.

Second, how has the distribution of birth spacing changed?
Although examiniong average birth intervals helps us understand the overall changes in 
spacing behavior, it tells us little about how these changes came about.
For example, the motivations that lead to an increase in average spacing are different if 
the increase comes mainly from parents trying to avoid very short birth intervals or from 
continued use of sex selection to ensure a son.

[TK this is unclear]
The traditional pattern of longer average spacing with sons appears to arise from 
an increase in the number of very long intervals, rather than from fewer very short 
birth intervals. 
Across education groups, there appears to be little difference in the likelihood of 
having a very short birth interval across sex compositions, especially for lower parities.

Third, to what extent does TFR reflect cohort fertility?
An advantage of the empirical model is that I can predict fertility behavior taking into
account both the likelihood of parity progression and having a child of a specific sex.
I compare those predictions with fertility rate estimates based on the same data to
show how the increases in birth spacing have led to a substantial overestimate of how 
fast fertility has declined. [TK too imprecise? underestimate of the level, rather?]

Finally, has the relationship between birth spacing and mortality changed over time?
[TK]

[TK results]

@@

Even as the level of female education has increased, the female labor force participation 
in both urban and rural areas has stagnated or decreased
\citep{Klasen2015,Afridi2018,Bhargava2018,Chatterjee2018,Bhargava2019}.
A decline in female labor force participation at the beginning of development is
consistent with the hypothesis of a U-shaped labor force participation as a country
develops \citep{Goldin1994}.
India's female labor force participation is, however, lower than most other
countries and more in line with countries in the Middle East and North Africa, and
does not yet show any signs of increasing
\citep{Klasen2015,Chatterjee2018}.

@@

>>>>

% Demography abstract
\noindent 
Strong son preference is typically associated with shorter birth spacing in the absence of 
sons, but access to sex selection has the potential to reverse this pattern because each 
abortion extends spacing by six to twelve months. 
I introduce an empirical method that simultaneously accounts for how sex selection 
increases both birth spacing and the likelihood of a son. 
Using the four rounds of India's National Family and Health Surveys, I show that birth 
intervals increased substantially over the last four decades, with the most significant
increases among women most likely to use sex selection.
Specifically, well-educated women with no boys now have significantly longer birth
intervals, and more male-biased sex ratios, than similar women with boys. 
Women with no education still follow the standard pattern of short average spacing when 
they have girls and little evidence of sex selection, with medium-educated women showing 
mixed results.
Furthermore, the traditional pattern of longer average spacing with sons appears to arise 
not from fewer very short birth intervals, but rather from the upper end of the distribution.

@@

Birth spacing has long served as a measure of son preference, with strong
son preference associated with shorter birth intervals the
fewer sons a family has \citep{ben-porath76b,Leung1988}.
However, the introduction of prenatal sex determination potentially 
changed the relationship between son preference and birth spacing.
The stronger the son preference, the more likely couples are to resort to sex-selective 
abortions, and each abortion increases the interval between births by six to twelve 
months.%
\footnote{
The increase consists of three parts.
First, starting from the abortion, the uterus needs at 
least two menstrual cycles to recover;  otherwise, the likelihood 
of spontaneous abortion increases substantially \citep{zhou00b}.
The second part is the waiting time to conception, which is between 
one and six months \citep{Wang2003}.
Finally, sex determination tests are reliable only from three months 
of gestation onwards.
}
Hence, longer birth intervals in the absence of sons, which previously would be taken as 
an indicator of lower son preference, may now instead arise because couples with a strong 
son preference use sex selection.

Here, I examine how birth spacing in India has changed over time and across 
groups with the introduction of sex selection.
I introduce an empirical method that directly incorporates the effects of 
sex-selective abortions on birth intervals
\emph{and} 
the likelihood of a son. 
The method can be used to analyze both situations with and without prenatal
sex selection.
I apply the method to birth histories of Hindu women, using data from the four
India's National Family and Health Surveys (NFHS), covering the period 
1972 to 2016.

% [Why India?]
India is a particularly compelling case with its long history of male preference, 
especially in the northern states \citep{Kishor1993,murthi95,arnold98}.
Mortality risk is higher for females than males for most ages groups, resulting in
an almost continuous increase in India's overall ratio of males to females 
over the last century \citep{dyson01,Navaneetham2011,Bongaarts2015}.
Differential stopping behavior is widespread, with an additional child more likely the
fewer sons a family has 
\citep{repetto72,Das1987,Arnold1997,arnold98,clark00,Basu2010,Barcellos2014}.%
\footnote{
Other countries show a similar pattern
\citep[see, for example,][]{larsen98,filmer09,Altindag2016}.
}
Furthermore, fertility has declined substantially, to the point where it is now at, or even 
below, replacement in some areas 
\citep{Guilmoto2013,Dharmalingam2014,International-Institute-for-Population-Sciences-IIPS2017}.
The combination of falling fertility, strong son preference, and expanding access to 
prenatal sex determination has lead to dramatic increases in the males-to-females ratio at 
birth over the last three decades
\citep{das_gupta97,Sudha1999,Arnold2002,retherford03b,jha06,Guilmoto2009a,Guilmoto2012,Bongaarts2013,Portner2015b,Jayachandran2017}.

% [Why spacing?] 
There are four main reasons for examining birth spacing and how it changes with 
access to sex selection.
First, researchers have made extensive use of birth spacing as a measure of son preference, 
and it is critical to understand to what extent spacing is still a useful measure of son 
preference.
For example, in Bangladesh, India, and Vietnam, having more boys, and particularly having 
a boy as the last-born child, significantly increase the duration to next birth,
\citep{Haughton1995,Haughton1996,Rahman1993,Bhalotra2008,Kumar2016,Soest2018}.%
\footnote{
Ethnic Indians in South Africa also show a longer duration after the birth of a 
son than after a daughter \citep{Gangadharan2003}.
}
Outside Asia, the evidence is more mixed, with North Africa showing shorter spacing in the 
absence of sons, while a similar pattern does not exist in Sub-Saharan Africa 
\citep{Rossi2015}.

Second, birth spacing can affect children's short and long-term outcomes.
Most of the research finds a negative impact of short spacing on child health, especially 
for very short intervals of 18 months or less \citep{Conde-Agudelo2006,Conde-Agudelo2012}.
There is evidence in countries as diverse as Bangladesh, Brazil, El Salvadore, India,
and Pakistan of worse health outcomes and a higher mortality risk the shorter the birth
interval
\citep{Cleland1984,Curtis1993,Whitworth2002,Bhargava2003,Rutstein2005,Bhalotra2008,Davanzo2008,Maitra2008,Makepeace2008,Gribble2009,Jayachandran2011,Saha2013,Jayachandran2017a,Ghosh2018}.
Most of the evidence on longer-term outcomes come from Western countries.
In the US, for example, a longer interval between births increases the older, but not the
younger, sibling's test scores \citep{Buckles2012}.
Furthermore, close spacing appears to increase the risk of dropping out of
high school and lower the probability of attending a post-secondary school in both the
US and Sweden \citep{Powell1993,Pettersson-Lidbom2009}.%
\footnote{
However, recent results from Sweden show no relationship between birth spacing, even if 
very short, and outcomes such as years of education completed, earnings, and unemployment 
\citep{Barclay2017}.
}

Third, the mother may also be affected by the length of the birth intervals, although 
reviews of the literature show mixed results on mother's anthropometric status 
\citep{Dewey2007,Conde-Agudelo2012}.
For India, there is, however, evidence that women with first-born girls are more likely to 
have anemia, possibly because short birth spacing can resulting in maternal depletion 
\citep{Milazzo2018}.
In Western countries, a longer interval between births has positive effects on long-term 
labor market outcomes, such as participation and income \citep{Gough2017,Karimi2014}.

Finally, we know less about what determines spacing behavior in developing countries than 
in developed countries, and with declining fertility and increasing numbers of women 
entering the labor force, understanding how couples make timing decisions is necessary 
for the design of suitable policies \citep{Portner2018}.%
\footnote{
It is clear that birth spacing does respond to policies in both developed and 
developing countries \citep{Pettersson-Lidbom2009,Todd2012,Meckel2015,Ghosh2018}.
}
Even the effects of access to contraceptive and declining fertility on spacing are unclear.
On the one hand, access to contraceptives allows women to avoid too short spacing between 
birth.
On the other hand, increased reliability of access and effectiveness of contraceptives can 
lead to shorter intervals between births if women used to have longer spacing to avoid
having too many children by accident \citep{Keyfitz1971,Heckman1976}.
These two counteracting effects may explain why better-educated women have shorter spacing 
than less educated women in some instances but not in others and why finding statistically 
significant effects of contraception use on birth spacing is difficult 
\citep{Tulasidhar1993,Whitworth2002,Bhalotra2008,Yeakey2009,Kim2010,Soest2018}.%
\footnote{
In Matlab, the effect of son preference on birth spacing is stronger in areas with better 
access to family planning \citep{Rahman1993}.
Other evidence, however, points to families ability to time births, even in the absence
of modern contraceptives \citep{Jayachandran2011,Alam2018}.
}

% [Hypotheses / questions]
Despite the increasing use of sex selection, the critical impacts of spacing on outcomes 
for both mother and children, and the extensive use of spacing to measure son preference, 
there has so far been no systematic examination of what effect access to sex selection 
have on spacing.
I address three questions to remedy this shortcoming in the literature.

First, how did average birth intervals change over time for different groups and how do 
these changes relate to the use of sex selection?
The hypothesis is that in the absence of sex selection son preference leads to shorter 
spacing the fewer boys a family has, whereas son preference increases spacing the fewer 
boys a family has when prenatal sex selection is available, possibly to the point where
we observe a reversal of the standard spacing pattern.
Families with son preference who---for one reason or another---do not use prenatal sex 
selection likely continue to have shorter birth spacing.
Based on the prior literature on sex selection, I focus on how birth intervals have
changed for different education levels, urban or rural, and the sex composition of 
prior children.

Second, have the shapes of the survival curves changed?
Although examining average birth intervals helps us understand the overall changes in 
spacing behavior, it tells us little about how these changes came about.
For example, the motivations that lead to an increase in average spacing are different if 
the increase comes mainly from parents trying to avoid very short birth intervals or from 
continued use of sex selection to ensure a son.

There are four main results.
First, the intervals between births have increased for all education groups over the four 
decades.
These increases are larger the higher the parity and the higher the education 
level.
Second, those women who are most likely to use sex selection---well-educated women with no
sons---have seen the largest increases in birth intervals.
As a result, well-educated women now show a reversal of the traditional spacing 
pattern, with the longest birth intervals for women with no sons.
Third, women who are least likely to use sex selection, those with no education in
rural areas, still follow the standard pattern of short spacing when they have girls and 
little evidence of the use of sex selection.
Finally, the traditional pattern of longer average spacing with sons appear to arise from 
more very long intervals, rather than from fewer very short birth intervals; across 
education groups, there appears to be little difference in the likelihood of having a very 
short birth interval across sex compositions, especially for lower parities.

@@

Even as the level of female education has increased, the female labor force participation 
in both urban and rural areas has stagnated or decreased
\citep{Klasen2015,Afridi2018,Bhargava2018,Chatterjee2018, Bhargava2019}.
A decline in female labor force participation at the beginning of development is
consistent with the hypothesis of a U-shaped labor force participation as a country
develops \citep{Goldin1994}.
India's female labor force participation is, however, lower than most other
countries and more in line with countries in the Middle East and North Africa, and
does not yet show any signs of increasing
\citep{Klasen2015,Chatterjee2018}.

@@

The prior literature show that shorter birth spacing, especially very short spacing,
is associated with higher risk of infant mortality.
Hence, it is possible that the increasing use of sex-selective abortions could
directly reduce infant mortality by lengthening the duration from the prior birth.

There are a number of potential issues with understanding to what extent this
has happened.

First, the sex of the next child is no longer random, and boys, in the absence of
differential treatment, have higher infant mortality than girls.

Second, it may be that what matters is not the time between births, but rather 
between pregnancies, in which case multiple abortions followed quickly by new 
pregnancies would lead to shorter real spells of recovery.
The implication is that long spells might have a higher mortality risk since
they are now more likely with multiple abortions.

Third, as I show above the use of sex selection is more prevalent the higher the
education level of the mother.
Higher maternal education is, however, also associated with lower mortality risk
to begin with.
On top of this, with economic growth and better health care, infant mortality
may have fallen across all education groups and areas in the period covered.

I address three questions.

First, does the increased use of sex selection lead to higher mortality risk?

Second, has the negative effects of short spacing changed over time?

Third, how do the effects of short spacing vary by education level in India?

There are a number of problems with estimating the relationship between 
sex-selective abortions and infant mortality.

First, there are potential selection problem.
For example, women who face higher mortality risk of their children might forego 
the use of sex selection to ensure that they reach their desired number of 
children.
If these women also know to avoid very short birth spacing, the precense of 
these women in the sample would inflate the mortality of the "middle" birth
spacing, relative to the very short or very long birth spacing.

[our JDE paper]

An additional selection with the opposite effect is that women who have
difficulties conceiving or carrying a preganancy to term may also have higher 
mortality risk for their offspring, leading to a supriuos correlation
between very long birth spacing and mortality \citep{Kozuki2013}. 
If better health care and higher income means that more of these women are
now able to carry their pregnancy to term we can end up with more births
after long spells that have higher mortality risk but not because of the
increased use of sex selection.

[Our Demography paper shows a different type of selection effect, where the
use of sex selection biases the effect estimates of other variables that
vary together with the use of sex selection.]

Given the data and that the focus of this paper is mainly on the birth
spacing, a more in-depth treatment of these issue is the topic of future work.
Part of the problem here is also that the method used previously, such
as family fixed effects does not work well when the number of births is
very low as it is for higher education women who use sex 
selection, especially since the spacing from marriage to first birth
does not provide any information \citep{Kozuki2013,Molitoris2019}.
However, the fixed effects results and linear probability do not deviate
substantially.

Second, the actual number of abortions are unobserved.
The results above do show, however, how the use of sex selection vary
by education and sex composition of prior children.
I, therefore, estimate infant mortality risk taking into account the
sex composition of the prior children, as well as the index child.

Starting with the sample used for estimating birth spacing, I select, by
parity, children born more than 12 months before the survey month.
The dependent variable is whether the child died within the first twelve
months of life (infant mortality).
As above, I restrict the analysis to parity two through four.

The main set of explanatory variable consists of dummies for birth spacing,
divided into less than 15 months from when conception could realistically
could happen (same definition as above) and based on the finding that 
spacing less than 24 months between birth is associated with higher
infant mortality \citep{Molitoris2019}, and then in 12 months intervals
until TK, where all births after and until 96 months are included.%
\footnote{
\citet{Molitoris2019} has used quartic functions, but the concern is
that with the relatively small number sample here the shape
of the function would be strongly influenced by a small number of
observations.
Using dummies provides a more flexible specification and the standard
error will directly reflect the higher degree of noise for the
spells periods with fewer observations.
}
In addition, I use dummies for sex of the index child (the child whose 
mortality is estimated) and the sex composition of the prior children.
Birth spacing dummies, sex of index child, and sex composition dummies
are all interacted.
The other explanatory variables are the same as above and estimations
are done separately by education level.

I estimate probability of infant mortality using a logit model.
Results are presented as the predicted probability of dying within the
first year at the possible combinations of index child sex, sex composition
of prior children, and the birth spacing dummied.
All other variables are set to their average values.
[TK ehh, did I also do urban/rural?]

Figures \ref{fig:mortality_low} to \ref{fig:mortality_highest} show the 
estimated probabilities.
To ease legibility all confidence intervals are suppressed
(graphs with confidence intervals are available on request).

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_low_dummies} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p1_low_dummies}
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_low_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p2_low_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_low_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p3_low_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_low_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p4_low_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_low}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_med_dummies} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p1_med_dummies}
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_med_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p2_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_med_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p3_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_med_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p4_med_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_med}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_high_dummies} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p1_high_dummies}
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_high_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p2_high_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_high_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p3_high_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_high_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p4_high_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_high}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_highest_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p2_highest_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_highest_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p3_highest_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_highest_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p4_highest_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_highest}
\end{figure}

For women with twelve or more years of education the results become very noisy
because of the low number of births, especially for parity three and four, and
the low mortality rate.
As above, I do not show results for the 1972--84 period because there were too
few women with this level of education for this period.

Based on the second parity births, women with twelve or more years of education
have the lowest infant mortality risk. 
Even for very short birth intervals the mortality risk is less than five percent,
and fall to less than half that for longer intervals.
Given the low mortality there is, not surprisingly, much difference between
the three periods covered.
Furthermore, there is no evidence that longer intervals are associated with
significantly higher mortality, especially for the groups most likely to have
used sex selection (women with a girl as their first child and a boy as their
second with a long interval).

Women with eight to eleven years of education have marginally higher mortality
risk than women with the most education, and also show slightly more evidence
of increased mortality with very short spacing.
Mortality risk have decreased slightly between periods, although the changes
are relatively minor.
As for women with the most education, there are not any statistically significant 
differences or increases in mortality with longer spacing.
Because of the low mortality risk and the relatively small number of births,
the estimates for the third and fourth parity are noisy.

Infant mortality risk for women with one to seven years of education have
clearly decreased over time.
This is also the first group that show increasing mortality risk with longer
spacing, although this effect mostly disappears for the later periods.
In other words, it does not appear to be the result of sex selection, but
rather arise from the selection bias discussed above where women with long
spells are those who find it difficult to either conceived and/or carry a 
pregnancy to term and also have higher mortality risk for their offspring.

Finally, women with no education have experienced substantial improvements
in infant mortality risk over time.
This decrease is especially pronounced for shorter intervals.
For the longer intervals, there is surprisingly little difference across
the education groups for the later periods.
